Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-29 colt-21/cvx-flats.tex
Checking over something in proof of Lem 27
Tweaking some statements for space and such

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (1) settings, such as structured prediction and extreme classification, the prediction dimension can easily become intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate\raf{cite}.
Challenges of type (2) include risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
Risk measures provably cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex: bounds for convex surrogates are left as a major open question.

@@

Various works have studied the minimum prediction dimension $d$ needed in order to construct a consistent surrogate loss $L: \reals^d \times \Y \to \reals$, typically through proxies such as calibration~\citep{steinwart2008support,agarwal2015consistent,ramaswamy2016convex} and property elicitation~\citep{frongillo2015vector-valued,fissler2016higher,frongillo2020elicitation}.
In Quadrant 1, \citet{ramaswamy2016convex} introduce a special case of convex consistency dimension (Definition~\ref{def:cvx-consistency-dim}), which has led to the design of consistent convex surrogates for discrete prediction problems such as hierarchical classification~\citep{ramaswamy2015hierarchical} and classification with an abstain option~\citep{ramaswamy2018consistent}.

@@

\begin{definition}[Convex Consistency Dimension]\label{def:cvx-consistency-dim}
  Given a target loss $\ell:\R \times\Y \to \reals$ or statistic $\gamma: \P \toto \R$, its \emph{convex consistency dimension} $\conscvx(\cdot)$ is the minimum dimension $d$ such that there is a $L \in \Lcvx_d$ and link $\psi$ such that $(L,\psi)$ is consistent with respect to $\ell$ or $\gamma$.
%	\btw{Maybe take out of definition environment, but (a) I'd worry that people wouldn't be able to find the definition when we use it later, and (b) I'm hoping we can make the distinction very clear, which now is achieved by the parallel definitions}
\end{definition}
A priori, this definition does not really make sense given a target statistic.
In Definition~\ref{def:consistent-prop}, we formalize consistency with respect to a property so that we can study $\conscvx(\gamma)$. \jessiet{Do we want to just talk about $\ell$ here?}\raft{I didn't think about the target property vs target loss point.  Let's discuss.}

@@

\begin{restatable}[\cite{ramaswamy2016convex} Theorem 18]{corollary}{hariresult}\label{cor:fsd-bound}
	Let $\ell:\R \times \Y \to \reals$ be a discrete loss eliciting $\gamma:\simplex \toto \R$ with $\Y$ finite.%, and let $L \in \Lcvx_d$ be consistent with respect to $\ell$.
	Then for all $p \in \simplex$ and $r \in \gamma(p)$,
	\begin{equation}
	  \conscvx(\gamma) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	  %d \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	\end{equation}
%% Bo: old version, Jun 4 7pm
%%	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
%%	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
%%	\begin{equation}
%%	\ccdim(\ell) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
%%	\end{equation}\jessiet{TODO: add minnability\bo{Tried!}}
\end{restatable}
\begin{proof}[Proof sketch]
	%\btw{Bo: rearranged the statements here a little, I felt the previous version required the reader to keep a bit more in their head.}
	First, take $p \in \relint{\simplex}$ for intuition, so $\|p\|_0 = n$.
	Lemma~\ref{lem:convex-flats-inf-dim} guarantees the existence of a $d$-flat $F$ with $p \in F \subseteq \gamma_r$.
	Moreover, since $p \in \relint{\simplex}$, the flat $F$ satisfies the requirements of Lemma~\ref{lem:finite-relint-dim}, and the bound $\dim(\Sc_{\gamma_r}(p)) \geq \dim(\Scr_F(p)) \geq n - d- 1$ is given by Lemma~\ref{lem:finite-relint-dim} (ii).
        \raft{should give intuition instead of (or in addition to) lemmas, since that's more the point of a sketch \jessie{Good point... taking a pass below}}

        \jessie{Adding an attempt at sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) \geq d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there is some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, we can verify $\Sc_{\gamma_r}(p) + \{p\}$ is one such $d$-flat.
    If $p \in \relint{\simplex}$, then the feasible subspace  (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-\dim(\Sc_{\gamma_r(p)})$.

@@

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width=.45\textwidth]{tikz/fsd-bound.pdf} &
		\includegraphics[width=.45\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
          (Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) geq 2$.}
	\label{fig:fsd-flats-abstain}
\end{figure}

@@

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) on the dimension of a consistent convex loss that is generally applicable and extends previous results from both the discrete (Corollary~\ref{cor:fsd-bound}) and continuous (Corollaries~\ref{cor:variance} and \ref{cor:spectral-risks}) estimation settings.

Several important questions remain open.
Particularly for the discrete settings, we would like to know whether one can lift the restriction that surrogates always achieve a minimum; we conjecture positively.
%For continuous settings, it would be especially interesting to extend to the case $\Y=\reals$, which would require a more careful analysis in Theorem~\ref{thm:bayes-risk-lower-bound}.\jessie{Flagging this bc we take care of it I hope}
Of course, we would like to characterize $\conscvx$ and $\eliccvx$ and develop a general framework for constructing surrogates achieving the best possible prediction dimension.
\jessie{for neurips, we had commented out everything below this for space}
One might be able to further tighten these bounds by property elicitation by studying monotonicity and adjacency of level sets.
In discrete predictions, these bounds might also be tightened if the equivalence of convex consistency dimension and embedding dimension of~\citet{finocchiaro2020embedding} is shown; the current embedding dimension bounds are not tight either, but additional structure is imposed by considering the embedding framework.
Moreover, the practical reason why consistency is desired is to ensure the guarantee of empirical risk minimization (ERM) rates; however, the relationship between ERM rates and property elicitation has not been studied.
%Lastly, Theorem~\ref{thm:cvx-flats} relies on Minkowski sums, which have not been clearly defined for infinite $\Y$; we leave this generalization as an open question.

@@

  We finally want to show $\dim(\affhull(\P)) = \dim(\spn(\P_p))$.
  Consider that any $q \in \spn(\P_p)$ can be written as a scalar multiple of an element of $\P_p$, which can be written as a convex combination of elements of the minimal basis $\P_p$.
  In particular, since $\vec 0 \in \P_p$, it can be written as an affine combination of elements of the basis, so $\dim(\affhull(\P)) \geq \dim(\spn(\P_p))$.
  We also have $\affhull(\P) - \{p\} \subseteq \spn(\P_p)$, so $\dim(\affhull(\P)) = \dim(\affhull(\P) - \{p\}) \leq \spn(\P_p)$.
  Therefore, $\dim(\affhull(\P)) = \dim(\spn(\P_p))$.
  \jessie{Check me!}

>>>>

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (1) settings, such as structured prediction and extreme classification, the prediction dimension can easily become intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate~\citep{ramaswamy2016convex}.
Challenges of type (2) include risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
Risk measures provably cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex: bounds for convex surrogates are left as a major open question.

@@

Various works have studied the minimum prediction dimension $d$ needed in order to construct a consistent surrogate loss $L: \reals^d \times \Y \to \reals$, typically through proxies such as calibration~\citep{steinwart2008support,agarwal2015consistent,ramaswamy2016convex} and property elicitation~\citep{frongillo2015vector-valued,fissler2016higher,frongillo2020elicitation}.
In Quadrant 1, \citet{ramaswamy2016convex} introduce a special case of convex consistency dimension (Definition~\ref{def:cvx-consistency-dim}), which led to consistent convex surrogates for discrete prediction problems such as hierarchical classification~\citep{ramaswamy2015hierarchical} and classification with an abstain option~\citep{ramaswamy2018consistent}.

@@

\begin{definition}[Convex Consistency Dimension]\label{def:cvx-consistency-dim}
  Given target loss $\ell:\R \times\Y \to \reals$ or statistic $\gamma: \P \toto \R$, its \emph{convex consistency dimension} $\conscvx(\cdot)$ is the minimum dimension $d$ such that $\exists L \in \Lcvx_d$ and link $\psi$ such that $(L,\psi)$ is consistent with respect to $\ell$ or $\gamma$.
%	\btw{Maybe take out of definition environment, but (a) I'd worry that people wouldn't be able to find the definition when we use it later, and (b) I'm hoping we can make the distinction very clear, which now is achieved by the parallel definitions}
\end{definition}
A priori, this definition does not really make sense given a target statistic.
In Definition~\ref{def:consistent-prop}, we formalize consistency with respect to a property so that we can study $\conscvx(\gamma)$. \jessiet{Do we want to just talk about $\ell$ here?}\raft{I didn't think about the target property vs target loss point.  Let's discuss.}

@@

\begin{restatable}[\cite{ramaswamy2016convex} Theorem 18]{corollary}{hariresult}\label{cor:fsd-bound}
	Let $\ell:\R \times \Y \to \reals$ be a discrete loss eliciting $\gamma:\simplex \toto \R$ with $\Y$ finite.%, and let $L \in \Lcvx_d$ be consistent with respect to $\ell$.
	Then for all $p \in \simplex$ and $r \in \gamma(p)$,
	\begin{equation}
	  \conscvx(\gamma) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	  %d \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	\end{equation}
%% Bo: old version, Jun 4 7pm
%%	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
%%	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
%%	\begin{equation}
%%	\ccdim(\ell) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
%%	\end{equation}\jessiet{TODO: add minnability\bo{Tried!}}
\end{restatable}
\begin{proof}[Proof sketch]
%	First, take $p \in \relint{\simplex}$ for intuition, so $\|p\|_0 = n$.
%	Lemma~\ref{lem:convex-flats-inf-dim} guarantees the existence of a $d$-flat $F$ with $p \in F \subseteq \gamma_r$.
%	Moreover, since $p \in \relint{\simplex}$, the flat $F$ satisfies the requirements of Lemma~\ref{lem:finite-relint-dim}, and the bound $\dim(\Sc_{\gamma_r}(p)) \geq \dim(\Scr_F(p)) \geq n - d- 1$ is given by Lemma~\ref{lem:finite-relint-dim} (ii).
        \raft{should give intuition instead of (or in addition to) lemmas, since that's more the point of a sketch \jessie{Good point... taking a pass below. Commented out original first paragraph above.}}

        \jessiet{Adding an attempt at first paragraph of sketch below, but also my brain isn't functioning.}
    If $\conscvx(\gamma) \geq d$, then there is a $L \in \Lcvx_d$ so that $L$ is consistent with respect to $\gamma$, and in turn, indirectly elicits $\gamma$.
    Lemma~\ref{lem:convex-flats-inf-dim} says that there is some $d$-flat $F = \zeros{V}$ contained in $\gamma_r$.
    In particular, we can verify $\Sc_{\gamma_r}(p) + \{p\}$ is one such $d$-flat.
    If $p \in \relint{\simplex}$, then the feasible subspace  (shifted by $p$) spans\jessiet{not the word I want...} $\relint{\simplex}$, and we can observe that the shifted feasible subspace is full dimensional relative to $\affhull(\simplex)$.
    We then use the rank-nullity theorem with a bound on $\rank(V) \leq d$ to obtain the bound $d \geq n-1-\dim(\Sc_{\gamma_r(p)})$.

@@

\begin{figure}[htb]
	\centering
	\begin{tabular}{cc}
		\includegraphics[width=.35\textwidth]{tikz/fsd-bound.pdf} &
		\includegraphics[width=.35\textwidth]{tikz/flats-bound.pdf} \\
	\end{tabular}
	\caption{(Left) Feasible subspace dimension $\Sc_{\gamma_\bot}(\bullet) = 2$ and $\Sc_{\gamma_\bot}({\color{blue}\star}) = 1$, giving the bound $\conscvx(\ell^{abs}) \geq 3- 1-1 = 1$.
          (Right) No $1$-flat through $\bullet$ (a line since $\bullet \in \relint{\simplex})$  stays fully contained in $\gamma_\bot$, so $\conscvx(\ell^{abs}) \geq 2$.}
	\label{fig:fsd-flats-abstain}
\end{figure}

@@

\section{Conclusions and future work}\label{sec:conclusions}
In this work, we show that indirect property elicitation can be a powerful necessary condition for the existence of a consistent surrogate loss (Theorem~\ref{thm:consistent-implies-indir-elic}).
Furthermore, we introduce a new lower bound (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) on convex consistency dimension that is generally applicable and extends previous results from both the discrete (Corollary~\ref{cor:fsd-bound}) and continuous (Corollaries~\ref{cor:variance} and \ref{cor:spectral-risks}) estimation settings.

Several important questions remain open.
Particularly for the discrete settings, we would like to know whether one can lift the restriction that surrogates always achieve a minimum; we conjecture positively.
%For continuous settings, it would be especially interesting to extend to the case $\Y=\reals$, which would require a more careful analysis in Theorem~\ref{thm:bayes-risk-lower-bound}.\jessie{Flagging this bc we take care of it I hope}
Of course, we would like to characterize $\conscvx$ and $\eliccvx$ and develop a general framework for constructing surrogates achieving the best possible prediction dimension.
Moreover, the practical reason why consistency is desired is to ensure the guarantee of empirical risk minimization (ERM) rates; however, the relationship between ERM rates and property elicitation has not been studied.
\btw{for neurips and colt, we had commented out everything below this for space}
%One might be able to further tighten these bounds by property elicitation by studying monotonicity and adjacency of level sets.
%In discrete predictions, these bounds might also be tightened if the equivalence of convex consistency dimension and embedding dimension of~\citet{finocchiaro2020embedding} is shown; the current embedding dimension bounds are not tight either, but additional structure is imposed by considering the embedding framework.

@@

  \jessiet{Check me; this is hand wavy, but I'm okay with it or submission!}  
  We finally want to show $\dim(\affhull(\P)) = \dim(\spn(\P_p))$.
  Consider that any $q \in \spn(\P_p)$ can be written as a scalar multiple of an element of $\P_p$, which can be written as a convex combination of elements of the minimal basis $\P_p$.
  In particular, since $\vec 0 \in \P_p$, it can be written as an affine combination of elements of the basis, so $\dim(\affhull(\P)) \geq \dim(\spn(\P_p))$.
  We also have $\affhull(\P) - \{p\} \subseteq \spn(\P_p)$, so $\dim(\affhull(\P)) = \dim(\affhull(\P) - \{p\}) \leq \spn(\P_p)$.
  Therefore, $\dim(\affhull(\P)) = \dim(\spn(\P_p))$.
